### 2022-11-15 NSF2022/Description/specificaim2.tex
all set: AMDG

% ----------------------------------------------------------------------
\begin{quotation}
  \noindent
  \textbf{Problem 5.} How can IEs be coupled with a Newton method to
  solve the non-linear PDE~\eqref{eqn:phase}?
\end{quotation}
\begin{wrapfigure}[18]{l}{2.3in}
  \vspace{-7pt}
  \centering
  \input{figures/doublewell.tikz} 
  \vspace{-10pt}
  \caption{\label{fig:CA} \footnotesize (a) A double-well potential. (b)
  The convergence of exact and inexact Newton's methods applied
  to~\eqref{eqn:phaseRoot}.}
\end{wrapfigure}
For a general double-well potential $W(\phi)$ with local minima at
$\phi_0$ and $\phi_1$, we propose to interpret the solution
of~\eqref{eqn:phase} as a root finding problem for
\begin{align}
  \label{eqn:phaseRoot}
  F[\phi] = -\epsilon^2 \Delta \phi + W'(\phi).
\end{align}
Equation~\eqref{eqn:phaseRoot} will be solved with Newton's method,
$\phi^{N+1} = \phi^{N} - \alpha (F'(\phi^N))^{-1} F(\phi^N)$, where
$F'(\phi^N) = -\epsilon^2 \Delta + W''(\phi^N)$ is the Fr\'{e}chet
derivative of $F$, and $\alpha$ is the step size that will be determined
using backtracking. This requires solving the variable coefficient PDE
$-\epsilon^2 \Delta v + W''(\phi^{N}) v = F[\phi^N]$, which is
challenging with an integral equation since a fundamental solution is
not readily available. However, by choosing a double-well potential with
$W''(\phi_0) = W''(\phi_1) = \beta > 0$, we propose to instead use an
inexact Newton's method where $(F')^{-1}$ is approximated with the
solution of $-\epsilon^2 \Delta v + \beta v = F[\phi^N]$. We test the
efficacy of the exact and inexact Newton's method applied
to~\eqref{eqn:phaseRoot} in $(0,1) \subset \RR$ by discretizing $F'$
with a second-order finite difference stencil. The double-well potential
is illustrated in Figure~\ref{fig:CA}(a), and both the exact and inexact
Newton's methods converge to the same solution. Figure~\ref{fig:CA}(b)
shows the norm of the residual at each Newton
iterate when using the exact second derivative (red curve) and the
approximate second derivative (blue curve). While the inexact Newton's
method requires roughly twice as many iterations to reach a tolerance
of $10^{-4}$, each iteration is significantly cheaper to apply.

>>>>

% ----------------------------------------------------------------------
\begin{quotation}
  \noindent
  \textbf{Problem 5.} How can IEs be coupled with a Newton method to
  solve the non-linear PDE~\eqref{eqn:phase}?
\end{quotation}
\begin{wrapfigure}[18]{l}{2.3in}
  \vspace{-7pt}
  \centering
  \input{figures/doublewell.tikz} 
  \vspace{-10pt}
  \caption{\label{fig:CA} \footnotesize (a) A double-well potential. (b)
  The convergence of exact and inexact Newton's methods applied
  to~\eqref{eqn:phaseRoot}.}
\end{wrapfigure}
For a general double-well potential $W(\phi)$ with local minima at
$\phi_0$ and $\phi_1$, we propose to interpret the solution
of~\eqref{eqn:phase} as a root finding problem for
\begin{align}
  \label{eqn:phaseRoot}
  F[\phi] = -\epsilon^2 \Delta \phi + W'(\phi).
\end{align}
Equation~\eqref{eqn:phaseRoot} will be solved with Newton's method,
$\phi^{N+1} = \phi^{N} - \xi (F'(\phi^N))^{-1} F(\phi^N)$, where
$F'(\phi^N) = -\epsilon^2 \Delta + W''(\phi^N)$ is the Fr\'{e}chet
derivative of $F$, and $\xi$ is the step size that will be determined
using backtracking. This requires solving the variable coefficient PDE
$-\epsilon^2 \Delta v + W''(\phi^{N}) v = F[\phi^N]$, which is
challenging with an integral equation since a fundamental solution is
not readily available. However, by choosing a double-well potential with
$W''(\phi_0) = W''(\phi_1) = \beta > 0$, we propose to instead use an
inexact Newton's method where $(F')^{-1}$ is approximated with the
solution of $-\epsilon^2 \Delta v + \beta v = F[\phi^N]$. We test the
efficacy of the exact and inexact Newton's method applied
to~\eqref{eqn:phaseRoot} in $(0,1) \subset \RR$ by discretizing $F'$
with a second-order finite difference stencil. The double-well potential
is illustrated in Figure~\ref{fig:CA}(a), and both the exact and inexact
Newton's methods converge to the same solution. Figure~\ref{fig:CA}(b)
shows the norm of the residual at each Newton
iterate when using the exact second derivative (red curve) and the
approximate second derivative (blue curve). While the inexact Newton's
method requires roughly twice as many iterations to reach a tolerance
of $10^{-4}$, each iteration is significantly cheaper to apply.
